---
title: "CRESULTS_sensory_predictors"
author: "Tara McAllister"
date: "2024-06-26"
output:
  html_document: default
  word_document: default
---

```{r chunk-setup, include=FALSE}
knitr::opts_chunk$set(include=TRUE, cache=TRUE, echo=FALSE, message=FALSE, warning=FALSE)
options(scipen = 999)
options(digits = 3)
```

```{r clear-environment}
########################################
#RUN THIS WITH CAUTION
########################################
rm(list=ls())
```

```{r package-load, echo=FALSE, include=FALSE, message=FALSE, warning=FALSE}
########################################
#LOAD PACKAGES
########################################
library("tidyverse")
library("ggpubr")
library("broom")
library("GGally")
library("flextable")

theme_set(theme_bw(base_size = 16))

#Create snapshot of package environment
# renv::init()
# renv::activate()
# renv::snapshot()
```


```{r define-helper-function-for-results-reporting}
#Formats a model coefficient as "beta = X.XX, SE = X.XX, p = X.XXX" for inline
#reporting in the text below. beta/SE are rounded to 2 digits and p to 3
#digits, matching the precision used throughout the manuscript.
coef_str <- function(mod, term, digits = 2) {
  x <- broom::tidy(mod)
  r <- x[x$term == term, ]
  if (nrow(r) != 1) {
    stop("coef_str: term '", term, "' matched ", nrow(r), " rows (expected 1). Check spelling/factor levels.")
  }
  p_txt <- ifelse(r$p.value < .001, "< .001",
                  paste0("= ", format(round(r$p.value, 3), nsmall = 2)))
  paste0(
    "β = ", round(r$estimate, digits),
    ", SE = ", round(r$std.error, digits),
    ", *p* ", p_txt
  )
}
```

```{r read-in-and-wrangle-data}
#Read in sensory data, cleaned for outliers
#Generated in CRESULTS_MOSAIC_sensory_wrangling.Rmd
#Variables starting with "new" have been standardized relative to the reference sample of typically developing children (age-binned z-scores)
#However, still need to reverse-code scores where a larger value represents worse performance, which is done here
sensorydata <- read.csv("RSSD_normed_sensory.csv") %>%
  dplyr::select(Subject = Study_ID, Age, Sex, normed_ident = newident,
                normed_discrim = newdiscrim,
                normed_catgood = newcatgood, normed_stereo = newstereo,
                normed_PAT = newPAT)

#Identification, discrimination, and stereognosis are reverse-coded here so
#that higher values uniformly reflect better performance across all five
#measures (category goodness and articulatory awareness already run in that
#direction and are left as-is)
sensorydata$normed_ident <- -1*sensorydata$normed_ident
sensorydata$normed_discrim <- -1*sensorydata$normed_discrim
sensorydata$normed_stereo <- -1*sensorydata$normed_stereo

#Read in summary outcome data (word probes)
#Generated in RCT_secondary_outcome_dosage.Rmd
outcomesdata <- read.csv("RCT_subjlevel_data.csv")

#Look at change over time
wide_data <- pivot_wider(outcomesdata,
                         id_cols = c(Subject, Condition,
                                     Randomization_Group,
                                     Baseline_Accuracy_Group),
                         names_from = Session,
                         values_from = mean_correct) %>%
  dplyr::arrange(Condition)
wide_data$change <- wide_data$POST_1 - wide_data$EVAL_1

#The sensory dataframe includes participants who did not complete the study;
#they are dropped in this merge
mydata <- left_join(wide_data, sensorydata, by = "Subject")

#Sanity check: accuracy by session and treatment condition
p1 <- ggplot(outcomesdata, aes(x = Session, y = mean_correct, fill = Condition))+
  geom_boxplot(notch=TRUE)
p1

#Age, gender breakdown
#na.rm=TRUE guards against a silent NA result if a future data update
#introduces missing Age values (none are present in the current dataset)
mean(mydata$Age, na.rm = TRUE)
sd(mydata$Age, na.rm = TRUE)
table(mydata$Sex)
table(mydata$Randomization_Group)

#Option to subset to biofeedback condition only
#mydata <- mydata[mydata$Randomization_Group!="Traditional",] %>% droplevels()

#Fix data types
mydata <- mydata %>%
  dplyr::mutate(across(where(is.character), as.factor))
```

```{r select-predictors-of-interest}
#Select only numeric columns, excluding the Subject ID itself so it doesn't
#show up as a spurious "predictor" in the correlation output below
df_numeric <- mydata %>% select(where(is.numeric)) %>% select(-Subject)

#Examine correlation of each numeric predictor with the outcome (change score)
cor_change <- cor(df_numeric, use = "pairwise.complete.obs")["change", ]
cor_change <- cor_change[names(cor_change) != "change"]

# Convert to tibble
cor_df <- tibble(predictor = names(cor_change), correlation = cor_change) %>%
   dplyr::arrange(desc(abs(correlation)))  # Sort by absolute correlation

# Count missing data
na_normed <- mydata %>%
   dplyr::summarise(across(starts_with("normed_"), ~ sum(is.na(.)))) %>%
  unlist()

# Plot as bar chart
p0 <- ggplot(cor_df, aes(x = reorder(predictor, abs(correlation)), y = correlation, fill = correlation)) +
  geom_col() +
  coord_flip() +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red") +
  labs(title = "Correlation of Predictors with Outcome",
       x = "Predictor", y = "Correlation") +
  theme_minimal()
p0

#Strongest somatosensory predictor is PAT; strongest auditory predictor is catgood
#(Note: this simple bivariate screen is exploratory only. The interaction
#models below retain all five sensory measures regardless of this ranking,
#per the independence rationale checked next.)
corA <- cor.test(mydata$change, mydata$normed_catgood)
corS <- cor.test(mydata$change, mydata$normed_PAT)

#Examine correlations between sensory values, to check whether it's
#reasonable to model the measures within each domain separately rather than
#pooling/averaging them
df_sensory <- mydata %>% select(starts_with("normed"))
cor_sensory <- round(cor(df_sensory, use = "pairwise.complete.obs"), digits = 2)
#Strongest positive pairwise correlation in the matrix (discrimination x identification)
cormax <- cor.test(mydata$normed_discrim, mydata$normed_ident)
#Strongest negative pairwise correlation in the matrix (category goodness x stereognosis)
corneg <- cor.test(mydata$normed_catgood, mydata$normed_stereo)
#ggpairs(df_sensory)
```

```{r export-correlation-matrix}
#Export the pairwise correlation matrix as Supplementary Table B
measure_labels <- c(normed_ident = "Identification", normed_discrim = "Discrimination",
                     normed_catgood = "Category Goodness", normed_stereo = "Stereognosis",
                     normed_PAT = "Articulatory Awareness")
vars <- names(measure_labels)

cor_sensory <- cor(df_sensory[vars], use = "pairwise.complete.obs")
p_sensory <- sapply(vars, function(a) sapply(vars, function(b)
  if (a == b) NA else cor.test(df_sensory[[a]], df_sensory[[b]])$p.value))

#Format each cell as "r.rr*" (p < .05) or "r.rr"; diagonal as an em dash
cor_display <- ifelse(row(cor_sensory) == col(cor_sensory), "\u2014",
                       paste0(formatC(cor_sensory, format = "f", digits = 2),
                              ifelse(p_sensory < .05, "*", "")))
dimnames(cor_display) <- list(measure_labels[vars], measure_labels[vars])

cor_table_B <- as.data.frame(cor_display) %>% tibble::rownames_to_column("Measure")

ft_corr <- flextable(cor_table_B)
ft_corr <- add_header_row(ft_corr, values = "Pairwise correlations among sensory measures (* p < .05)", colwidths = ncol(cor_table_B))
ft_corr <- autofit(ft_corr)
save_as_docx(ft_corr, path = "Supplement_Table_B.docx")
```

The exclusion of corrupted data and outliers resulted in the exclusion of the identification task from n = `r na_normed["normed_ident"]` participants, the discrimination task from n = `r na_normed["normed_discrim"]` participants, the category goodness judgment task from n = `r na_normed["normed_catgood"]` participants, and the articulatory awareness task from n = `r na_normed["normed_PAT"]` participant. In addition, the stereognosis task represented a special case where data were missing from a large number of participants due to sanitary precautions during the COVID-19 pandemic; this resulted in missing data from n = `r na_normed["normed_stereo"]` participants. Given that the sources of missingness appear to be largely random with respect to our treatment groups, we anticipate that the exclusion of these data will not bias our results.


```{r stats-without-pooling}
#Correlations (above) indicate that the sensory measures within each domain
#are only weakly correlated with one another, suggesting each captures a partially
#independent facet of that domain. This is why the models below retain the
#individual auditory and somatosensory measures rather than pooling/averaging
#across them.

#Treatment condition interaction with auditory acuity
mydata$Randomization_Group <- relevel(mydata$Randomization_Group, ref = "Ultrasound")
modA <- lm(change ~ Randomization_Group * (normed_catgood + normed_discrim + normed_ident), data = mydata)
#Alternative model with control for baseline accuracy
altmodA <- lm(POST_1 ~ Randomization_Group * (normed_catgood + normed_discrim + normed_ident) + EVAL_1, data = mydata)
#summary(modA)
#summary(altmodA)

#Relevel to inspect traditional MBT and VAB
#Including alternative models with control for baseline accuracy
mydata$Randomization_Group <- relevel(mydata$Randomization_Group, ref = "Traditional")
modArelT <- lm(change ~ Randomization_Group * (normed_catgood + normed_discrim + normed_ident), data = mydata)
altmodArelT <- lm(POST_1 ~ Randomization_Group * (normed_catgood + normed_discrim + normed_ident) + EVAL_1, data = mydata)
mydata$Randomization_Group <- relevel(mydata$Randomization_Group, ref = "Visual Acoustic")
modArelV <- lm(change ~ Randomization_Group * (normed_catgood + normed_discrim + normed_ident), data = mydata)
altmodArelV <- lm(POST_1 ~ Randomization_Group * (normed_catgood + normed_discrim + normed_ident) + EVAL_1, data = mydata)

#Biofeedback interaction with somatosensory acuity
mydata$Randomization_Group <- relevel(mydata$Randomization_Group, ref = "Ultrasound")
modS <- lm(change ~ Randomization_Group * (normed_PAT + normed_stereo), data = mydata)
#Alternative model with control for baseline accuracy
altmodS <- lm(POST_1 ~ Randomization_Group * (normed_PAT + normed_stereo) + EVAL_1, data = mydata)
#summary(modS)
#summary(altmodS)

#Relevel to inspect traditional MBT and VAB
#Including alternative models with control for baseline accuracy
mydata$Randomization_Group <- relevel(mydata$Randomization_Group, ref = "Traditional")
modSrelT <- lm(change ~ Randomization_Group * (normed_PAT + normed_stereo), data = mydata)
altmodSrelT <- lm(POST_1 ~ Randomization_Group * (normed_PAT + normed_stereo) + EVAL_1, data = mydata)
mydata$Randomization_Group <- relevel(mydata$Randomization_Group, ref = "Visual Acoustic")
modSrelV <- lm(change ~ Randomization_Group * (normed_PAT + normed_stereo), data = mydata)
altmodSrelV <- lm(POST_1 ~ Randomization_Group * (normed_PAT + normed_stereo) + EVAL_1, data = mydata)

#Per-arm n's for the somatosensory complete-case subsample — related to 
#balance-across-arms caveat noted in the Discussion/abstract
mydata_stereo_cc <- mydata %>%
  filter(!is.na(normed_stereo))
table(mydata_stereo_cc$Randomization_Group)
```

In the model examining auditory predictors (n = `r length(modA$residuals)`), there was a positive direction of association between discrimination score and change in accuracy in the ultrasound biofeedback treatment condition, although it fell just short of the conventional threshold for statistical significance (`r coef_str(modA, "normed_discrim")`). There was no significant association between discrimination score and change in accuracy for the other two treatment conditions. There was a significant interaction between discrimination score and treatment group (visual-acoustic versus ultrasound: `r coef_str(modA, "Randomization_GroupVisual Acoustic:normed_discrim")`). Visual inspection of Figure 2 shows that the slope of the association for visual-acoustic biofeedback was negative, such that children with lower discrimination scores tended to exhibit a larger degree of change over the course of treatment than children with higher discrimination scores. By contrast, MBT and ultrasound biofeedback showed a positive relationship between discrimination score and change over the course of treatment. There were no significant effects or interactions for the other auditory measures (identification and category goodness judgment); see Supplementary Table A for full results.

```{r model-summary, cache = TRUE}
#Generate supplementary document reporting full results of regressions

models <- list(
  "Auditory, ref = Ultrasound" = modA,
  "Auditory, ref = MBT" = modArelT,
  "Auditory, ref = Visual-Acoustic" = modArelV,
  "Somatosensory, ref = Ultrasound" = modS,
  "Somatosensory, ref = MBT" = modSrelT,
  "Somatosensory, ref = Visual-Acoustic" = modSrelV,
  "Auditory controlling for baseline, ref = Ultrasound" = altmodA,
  "Auditory controlling for baseline, ref = MBT" = altmodArelT,
  "Auditory controlling for baseline, ref = Visual-Acoustic" = altmodArelV,
  "Somatosensory controlling for baseline, ref = Ultrasound" = altmodS,
  "Somatosensory controlling for baseline, ref = MBT" = altmodSrelT,
  "Somatosensory controlling for baseline, ref = Visual-Acoustic" = altmodSrelV
)


tables <- lapply(names(models), function(nm) {
  df <- broom::tidy(models[[nm]]) |>
    mutate(
      estimate  = formatC(estimate, format = "f", digits = 2),
      std.error = formatC(std.error, format = "f", digits = 2),
      statistic = formatC(statistic, format = "f", digits = 2),
      p.value   = formatC(p.value, format = "f", digits = 3)
    )
  ft <- flextable(df)
  ft <- add_header_row(ft, values = nm, colwidths = ncol(df))
  ft <- autofit(ft)
  ft
})
names(tables) <- names(models)

do.call(save_as_docx, c(tables, path = "Supplement_Model_Results.docx"))
```


```{r inspect-interaction-auditory}
#With discrim

mydata$Randomization_Group <- factor(
  mydata$Randomization_Group,
  levels = c("Traditional", "Visual Acoustic", "Ultrasound"),
  labels = c("MBT", "Visual Acoustic", "Ultrasound")
)


#Define palettes
pal <- c("#54278F", "#2B8CBE", "#9E9AC8")
shape_pal <- c(21, 22, 24)

pAd <- ggplot(
  mydata,
  aes(
    x = normed_discrim,
    y = change,
    color = Randomization_Group,
    fill  = Randomization_Group,
    shape = Randomization_Group
  )
) +
  geom_point(size = 1.5, stroke = 1) +
  geom_smooth(
    method = "lm",
    se = TRUE,
    alpha = 0.35,      # higher than before
    linewidth = 1.2
  ) +
  scale_color_manual(values = pal) +
  scale_fill_manual(values  = pal) +
  scale_shape_manual(values = shape_pal) +
  theme_bw(base_size = 16) +
  ylab("Change in accuracy") +
  xlab("Discrimination score (age-normalized)") +
  labs(color = "Group", fill = "Group", shape = "Group") +
  theme(
    legend.position = c(.25, .75),
    legend.justification = c("right", "bottom"),
    legend.background = element_rect(fill = "transparent", color = NA)
  )

pAd
```


In the model examining somatosensory predictors (n = `r length(modS$residuals)`), there was a positive association between stereognosis score and change in accuracy in the visual-acoustic biofeedback treatment condition (`r coef_str(modSrelV, "normed_stereo")`), but not in the other conditions. There were significant interactions between age-normalized stereognosis score and treatment group (visual-acoustic versus MBT: `r coef_str(modSrelT, "Randomization_GroupVisual Acoustic:normed_stereo")`; visual-acoustic versus ultrasound: `r coef_str(modS, "Randomization_GroupVisual Acoustic:normed_stereo")`). Visual inspection of Figure 3 reveals that MBT and ultrasound biofeedback showed a slightly negative relationship between stereognosis score and change over the course of treatment, such that children with lower stereognosis scores tended to exhibit a larger treatment response than children with higher scores. This contrasted with the positive direction of association between stereognosis score and response to treatment in the visual-acoustic biofeedback condition.


```{r inspect-interaction-somatosensory}
#With stereognosis

pSs <- ggplot(
  mydata,
  aes(
    x = normed_stereo,
    y = change,
    color = Randomization_Group,
    fill  = Randomization_Group,
    shape = Randomization_Group
  )
) +
  geom_point(size = 1.5, stroke = 1) +
  geom_smooth(
    method = "lm",
    se = TRUE,
    alpha = 0.35,
    linewidth = 1.2
  ) +
  scale_color_manual(values = pal) +
  scale_fill_manual(values  = pal) +
  scale_shape_manual(values = shape_pal) +
  ylab("Change in accuracy") +
  xlab("Stereognosis score (age-normalized)") +
  theme_bw(base_size = 16) +
  labs(color = "Group", fill = "Group", shape = "Group") +
  theme(
    legend.position = c(0.97, 0.03),  # bottom right
    legend.justification = c("right", "bottom"),
    legend.background = element_rect(fill = "transparent", color = NA)
  )

pSs
```
